---
title: "Assignment 7"
author: "Adrian Bracher (Matr. Nr. 01637180)"
date: "12.05.2021"
output:
  pdf_document: 
    toc: true 
    toc_depth: 3
    number_sections: true 
---

# Introduction to Factor Variables
```{r echo=FALSE, results='hide'}
multiple_choice_responses = read.csv("smc_with_js.csv")
library(tidyverse)
library(dplyr)
flying_etiquette = read.csv("flying-etiquette.csv")
```
## Recognizing factor variables
We start this chapter by simply printing out the data frame and then checking if a column of the data frame contains data of type factor.
```{r echo = T, results='hide'}
# Print out the dataset
print(multiple_choice_responses)

# Check if CurrentJobTitleSelect is a factor
is.factor(multiple_choice_responses$CurrentJobTitleSelect)
```
## Getting number of levels
The tidyverse functions mutate_if(), summarise_all() and gather() are introduced in this exercise. The comments describe how they are applied in the code below.
```{r}
# Change all the character columns to factors
responses_as_factors <- multiple_choice_responses %>%
    mutate_if(is.character, as.factor)

number_of_levels <- responses_as_factors %>%
	# apply the function nlevels to each column
    summarise_all(nlevels) %>%
    # change the dataset from wide to long
    gather(variable, num_levels)
```

## Examining number of levels
In the code below we use top_n to select the 3 rows with the highest number of levels. pull() is used to extract a column and removing the name, which leaves the values from the column.
```{r}
# Select the 3 rows with the highest number of levels
number_of_levels %>%
    top_n(3, num_levels)
    
number_of_levels %>%
	# filter for where the column called variable equals CurrentJobTitleSelect
    filter(variable == "CurrentJobTitleSelect") %>%
	# pull num_levels
    pull(num_levels)
```
## Examining levels
In this exercise we call the function levels() to get all (distinct) levels of the input column.
```{r}
responses_as_factors %>%
    # pull CurrentJobTitleSelect
    pull(CurrentJobTitleSelect) %>%
    # get the values of the levels
    levels()
```
## Reordering a variable by its frequency
The package "forcats" allows us to modify ggplots in various ways. Here, we first flip the coordinates with "+ coord_flip()" and call fct_infreq() to order the data and then fct_rev() to reverse the order.
```{r}
# Make a bar plot
ggplot(multiple_choice_responses, aes(x=fct_rev(fct_infreq(EmployerIndustry)))) + 
    geom_bar() + 
    # flip the coordinates
    coord_flip()
```

## Ordering one variable by another
As the final exercise in this chapter we create a plot. We use mutate(), in combination with fct_reorder() to reorder EmployerIndustry and add a column mean_age with summarise(). 
```{r}
multiple_choice_responses %>%
  # remove NAs
  filter(!is.na(EmployerIndustry) & !is.na(Age)) %>%
  # get mean_age by EmployerIndustry
  group_by(EmployerIndustry) %>%
  summarise(mean_age = mean(Age)) %>%
  # reorder EmployerIndustry by mean_age 
  mutate(EmployerIndustry = fct_reorder(EmployerIndustry, mean_age)) %>%
  # make a scatterplot of EmployerIndustry by mean_age
  ggplot(aes(x = EmployerIndustry, y = mean_age)) + 
    geom_point() + 
    coord_flip()
```
# Manipulating Factor Variables
## Changing the order of factor levels
Here we learn how to use fct_relevel to manually change the order of factor levels. Then, we make a bar plot of the results.
```{r echo=FALSE, results='hide'}
multiple_choice_responses = read.csv("multiple_choice_responses-2.csv")
```

```{r}
# Get the levels of WorkInternalVsExternalTools
levels(multiple_choice_responses$WorkInternalVsExternalTools)

# Reorder the levels from internal to external 
mc_responses_reordered <- multiple_choice_responses %>%
    mutate(WorkInternalVsExternalTools = fct_relevel(WorkInternalVsExternalTools,"Entirely internal", "More internal than external", "Approximately half internal and half external", "More external than internal", "Entirely external", "Do not know"))
# Make a bar plot of the responses
ggplot(mc_responses_reordered, aes(x=WorkInternalVsExternalTools)) + 
    geom_bar() + 
    coord_flip()
```
## Tricks of fct_relevel()
The reordering of levels with fct_relevel() can take very advanced forms with multiple parameters, for example last = Inf puts the specified level at the very last.
```{r echo=FALSE, results='hide'}
multiple_choice_responses = read.csv("multiple_choice_responses-3.csv")
```

```{r}
multiple_choice_responses %>%
    # Move "I did not complete any formal education past high school" and "Some college/university study without earning a bachelor's degree" to the front
    mutate(FormalEducation = fct_relevel(FormalEducation, "I did not complete any formal education past high school", "Some college/university study without earning a bachelor's degree")) %>%
    # Move "I prefer not to answer" to be the last level.
    mutate(FormalEducation = fct_relevel(FormalEducation, "I prefer not to answer", after = Inf)) %>%
    # Move "Doctoral degree" to be after the 5th level
    mutate(FormalEducation = fct_relevel(FormalEducation, "Doctoral degree", after=5)) %>%
    # Examine the new level order
    pull(FormalEducation) %>%
    levels()
```

## Renaming a few levels
In the following we decide to shorten the value of some levels with fct_recode().
```{r}
multiple_choice_responses %>%
    # rename the appropriate levels to "High school" and "Some college"
    mutate(FormalEducation = fct_recode(FormalEducation,
    "High school" = "I did not complete any formal education past high school", 
    "Some college" = "Some college/university study without earning a bachelor's degree")) %>%
    # make a bar plot of FormalEducation
    ggplot(aes(x = FormalEducation)) + 
    geom_bar()
```
## Manually collapsing levels
In this exercise we learn how to collapse multiple levels into one with fct_collapse() and then also collapse all the others, unspecified, into one level "Other" with fct_other().
```{r}
multiple_choice_responses %>%
    # Create new variable, grouped_titles, by collapsing levels in CurrentJobTitleSelect
    mutate(grouped_titles = fct_collapse(CurrentJobTitleSelect, 
        "Computer Scientist" = c("Programmer", "Software Developer/Software Engineer"), 
        "Researcher" = "Scientist/Researcher", 
        "Data Analyst/Scientist/Engineer" = c("DBA/Database Engineer", "Data Scientist", 
                                              "Business Analyst", "Data Analyst", 
                                              "Data Miner", "Predictive Modeler"))) %>%
    # Keep all the new titles and turn every other title into "Other"
    mutate(grouped_titles = fct_other(grouped_titles, keep = c("Computer Scientist", "Researcher", "Data Analyst/Scientist/Engineer"))) %>% 
    # Get a count of the grouped titles
    count(grouped_titles)
```
## Lumping variables by proportion
The function fct_lump can be used for collapsing the least common levels into "other". 
```{r}
multiple_choice_responses %>%
  # remove NAs of MLMethodNextYearSelect
  filter(!is.na(MLMethodNextYearSelect)) %>%
  # create ml_method, which lumps all those with less than 5% of people into "Other"
  mutate(ml_method = fct_lump(MLMethodNextYearSelect, prop=0.05)) %>%
  # count the frequency of your new variable, sorted in descending order
  count(ml_method, sort=TRUE)
```
## Preserving the most common levels
This time we don't collapse a fixed proportion, but all but a fixed amount of levels (in this case 5).
```{r}
multiple_choice_responses %>%
  # remove NAs 
  filter(!is.na(MLMethodNextYearSelect)) %>%
  # create ml_method, retaining the 5 most common methods and renaming others "other method" 
  mutate(ml_method = fct_lump(MLMethodNextYearSelect, n=5, other_level = "other method")) %>%
  # count the frequency of your new variable, sorted in descending order
  count(ml_method, sort=TRUE)
```
# Creating Factor Variables
## Grouping and reshaping similar columns
In this exercise, we learn how to use str_remove in combination with mutate.
```{r}
learning_platform_usefulness <- multiple_choice_responses %>%
  # select columns with LearningPlatformUsefulness in title
  select(contains("LearningPlatformUsefulness")) %>%
  # change data from wide to long
  gather(learning_platform, usefulness) %>%
  # remove rows where usefulness is NA
  filter(!is.na(usefulness)) %>%
  # remove "LearningPlatformUsefulness" from each string in learning_platform 
  mutate(learning_platform = str_remove(learning_platform, "LearningPlatformUsefulness"))
```
## Summarizing data
In the code below we create a faceted plot with the dplyr function add_count().
```{r}
perc_useful_platform <- learning_platform_usefulness %>%
  # change dataset to one row per learning_platform usefulness pair with number of entries for each
  count(learning_platform, usefulness) %>%
  # use add_count to create column with total number of answers for that learning_platform
  add_count(learning_platform, wt = n, name='nn') %>%
  # create a new column, perc, that is the percentage of people giving that response for that learning_platform
  mutate(perc = n / nn)

# create a line graph for each question with usefulness on x-axis and percentage of responses on y
ggplot(perc_useful_platform, aes(x = usefulness, y = perc, group = learning_platform)) + 
  geom_line() + 
  facet_wrap(~ learning_platform)
```

## Creating an initial plot
In this exercise we create a plot using various dplyr functions, and the function if_else() is introduced.
```{r}
usefulness_by_platform <- learning_platform_usefulness %>%
    # If usefulness is "Not Useful", make 0, else 1 
    mutate(usefulness = if_else(usefulness == "Not Useful", 0, 1)) %>%
    # Group by learning platform 
    group_by(learning_platform) %>%
    # Summarize the mean usefulness for each platform
    summarise(avg_usefulness = mean(usefulness))
# Make a scatter plot of average usefulness by learning platform 
ggplot(usefulness_by_platform, aes(x=learning_platform, y=avg_usefulness)) + 
        geom_point()
```
## Editing plot text
In this exercise we learn how to rotate x-axis test, rename the axis labels and change the axis scale.
```{r}
ggplot(usefulness_by_platform, aes(x = learning_platform, y = avg_usefulness)) + 
    geom_point() + 
    # rotate x-axis text by 90 degrees
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    # rename y and x axis labels
    labs(x = "Learning Platform", y = "Percent finding at least somewhat useful") + 
    # change y axis scale to percentage
    scale_y_continuous(labels = scales::percent)
```

## Reordering graphs
Here we learn how to apply fct_reorder and fct_rev inside a mutate call.
```{r}
usefulness_by_platform %>%
	# reorder learning_platform by avg_usefulness
	mutate(learning_platform = fct_reorder(learning_platform, avg_usefulness)) %>%
	# reverse the order of learning_platform
	mutate(learning_platform = fct_rev(learning_platform)) %>%
	ggplot(aes(x = learning_platform, y = avg_usefulness)) + 
	geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    labs(x = "Learning Platform", y = "Percent finding at least somewhat useful") + 
    scale_y_continuous(labels = scales::percent)
```
## case_when() with single variable
In this exercise we learn how to use the function case_when().
```{r echo=FALSE, results='hide'}
multiple_choice_responses = read.csv("multiple_choice_responses-4.csv")
```

```{r}
# Check the min age
min(multiple_choice_responses$Age, na.rm=TRUE)

# Check the max age
max(multiple_choice_responses$Age, na.rm=TRUE)

multiple_choice_responses %>%
    # Filter for rows where Age is between 10 and 90
    filter(between(Age, 10, 90)) %>%
    # Create the generation variable based on age
    mutate(generation = case_when(
    between(Age, 10, 22) ~ "Gen Z", 
    between(Age, 23, 37) ~ "Gen Y", 
    between(Age, 38, 52) ~ "Gen X", 
    between(Age, 53, 71) ~ "Baby Boomer", 
    between(Age, 72, 90) ~ "Silent"
    )) %>%
    # Get a count of how many answers in each generation
    count(generation)
```
## case_when() from multiple columns
We extend our knowledge from the previous exercise using case_when to multiple columns. 
```{r echo=FALSE, results='hide'}
multiple_choice_responses = read.csv("multiple_choice_responses-5.csv")
```

```{r}
multiple_choice_responses %>%
    # Filter out people who selected Data Scientist as their Job Title
    filter(CurrentJobTitleSelect != "Data Scientist") %>%
    # Create a new variable, job_identity
    mutate(job_identity = case_when(
        CurrentJobTitleSelect == "Data Analyst" & 
        DataScienceIdentitySelect == "Yes" ~ "DS analysts", 
        CurrentJobTitleSelect == "Data Analyst" & 
        DataScienceIdentitySelect %in% c("No", "Sort of (Explain more)") ~ "NDS analyst", 
        CurrentJobTitleSelect != "Data Analyst" & 
        DataScienceIdentitySelect == "Yes" ~ "DS non-analysts", 
        TRUE ~ "NDS non analysts")) %>%
    # Get the average job satisfaction by job_identity, removing NAs
    group_by(job_identity) %>%
    summarise(avg_js = mean(JobSatisfaction, na.rm=TRUE))
```
# Case Study on Flight Etiquette
## Changing characters to factors
In the code below we use mutate_if to convert all characters to factors
```{r results='hide'}

flying_etiquette %>%
    # Change characters to factors
    mutate_if(is.character, as.factor) %>%
    # Filter out those who have never flown on a plane
    filter('How often do you travel by plane?' != "Never")
```

## Tidying data
We apply select, contains, gather and filter to the flying_etiquette data frame in order to tidy them.
```{r}
gathered_data <- flying_etiquette %>%
    mutate_if(is.character, as.factor) %>%
    filter('How often do you travel by plane?' != "Never") %>%
    # Select columns containing "rude"
    select(contains("rude")) %>%
    # Change format from wide to long
    gather(response_var, value)
```

## Cleaning up strings
The function str_remove can also be used with regex expressions like ".*" as you can see in the following:
```{r  results='hide'}
gathered_data %>%
    # Remove everything before and including "rude to " (with that space at the end!)
    mutate(response_var = str_remove(response_var, ".*rude to ")) %>%
    # Remove "on a plane"
    mutate(response_var = str_remove(response_var, "on a plane"))
```
## Dichotomizing variables
Here we use if_else, str_replace and mutate to make a new column called rude, which contains a binary value 1 or 0 to distinguish rudeness.
```{r}
dichotimized_data <- gathered_data %>%
    mutate(response_var = str_replace(response_var, '.*rude to ', '')) %>%
    mutate(response_var = str_replace(response_var, 'on a plane', '')) %>%
	# Remove rows that are NA in the value column
	filter(!is.na(value)) %>%
    # Dichotomize the value variable to make a new variable, rude
    mutate(rude = if_else(value %in% c('No, not rude at all', 'No, not at all rude'), 0, 1))
```
## Summarizing data
Again, we apply summarise to add a new column perc_rude that aggregates another column.
```{r}
rude_behaviors <- gathered_data %>%
    mutate(response_var = str_replace(response_var, '.*rude to ', '')) %>%
    mutate(response_var = str_replace(response_var, 'on a plane', '')) %>%
	# Remove rows that are NA in the value column
	filter(!is.na(value)) %>%
    mutate(rude = if_else(value %in% c("No, not rude at all", "No, not at all rude"), 0, 1)) %>%
    # Group by response_var
    group_by(response_var) %>%
    # Create perc_rude, the percent considering each behavior rude
    summarise(perc_rude = mean(rude))

rude_behaviors
```
## Creating an initial plot
We create a simple plot using functions such as fct_reorder.
```{r}
initial_plot <- rude_behaviors %>%
	# reorder response_var by perc_rude
	mutate(response_var = fct_reorder(response_var, perc_rude)) %>%
	# make a bar plot of perc_rude by response_var
	ggplot(aes(x = response_var, y = perc_rude)) + 
    geom_col()


# View your plot
initial_plot
```
## Fixing labels
We use the labs function to change title, subtitle, caption, x -axis and y-axis labels.
```{r}
titled_plot <- initial_plot + 
    # Add the title, subtitle, and caption
    labs(title = "Hell Is Other People In A Pressurized Metal Tube",
         subtitle = "Percentage of 874 air-passenger respondents who said action is very or somewhat rude",
         caption = "Source: SurveyMonkey Audience", 
         # Remove the x- and y-axis labels
         x = "", 
         y = "") 

titled_plot
```
## Flipping things around
In this exercise we use theme with axis.text as well as axis.ticks to change the plot to our wishes.
```{r}
flipped_plot <- titled_plot + 
    # Flip the axes
    coord_flip() + 
    # Remove the x-axis ticks and labels
    theme(axis.text.x = element_blank(), 
        axis.ticks.x = element_blank())
```
## Finalizing the chart
We set the label parameter to percent(perc_rude) to label each bar with the percent rude value.
```{r eval=FALSE}
library(memisc)
flipped_plot + 
    # Apply percent() to perc_rude to label above the bar with the perc value
    geom_text(aes(label = percent(perc_rude), 
                  y = perc_rude + .03), 
              position = position_dodge(0.9),
              vjust = 1)
```
