---
title: "Assignment 1"
author: "Adrian Bracher"
output: pdf_document
---

# Importing data from flat files with utils

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A comma-separated values (csv) file can be read and converted to a data frame by calling read.csv with the file path as parameter.
The function call str(pools) prints out the structure of the data frame pools that we just created.

```{r}
# Import swimming_pools.csv: 
pools <- read.csv("swimming_pools.csv")


# Print the structure of pools
str(pools)
```

Since we do not want to import the data entries as factors, stringsAsFactors has to be specified as well.

```{r}
# Import swimming_pools.csv correctly: pools
pools <- read.csv("swimming_pools.csv", stringsAsFactors = FALSE)

# Check the structure of pools
str(pools)
```

The function read.delim is similar to read.csv, but used for tab separated values instead of csv. The parameter header specifies if the first line in the input consists of a header or data.
Summary prints summary statistics about all variables in the specified data frame.
```{r}
# Import hotdogs.txt: 
hotdogs <- read.delim("hotdogs.txt", header = FALSE)
```

The function read.table has to be used for reading in data that is neither csv nor tab separated. Instead a separator can be specified with the parameter sep. The parameter col.names adds header information to the dataframe.
head(hotdogs) prints out the first couple of observations of hotdogs data frame.
```{r}


# Import the hotdogs.txt file: hotdogs
hotdogs <- read.table("hotdogs.txt", 
                      sep = "\t", 
                      col.names = c("type", "calories", "sodium"))

# Call head() on hotdogs
head(hotdogs)
```

In this exercise the read.delim function had to be completed, header is set to false, because hotdogs.txt does not contain header information and the header is instead provided with the col.names parameter. which.min and which.max return the index of the observation with the lowest and highest value in the given variable.
```{r}
# Finish the read.delim() call
hotdogs <- read.delim("hotdogs.txt", header = FALSE, col.names = c("type", "calories", "sodium"))

# Select the hot dog with the least calories: lily
lily <- hotdogs[which.min(hotdogs$calories), ]

# Select the observation with the most sodium: tom
tom <- hotdogs[which.max(hotdogs$sodium), ]

# Print lily and tom
lily
tom
```

The parameter colClasses can be used to explicitly set the classes of the columns. If a column class is set to NULL, then the column will not be imported.
```{r}
hotdogs2 <- read.delim("hotdogs.txt", header = FALSE, 
                       col.names = c("type", "calories", "sodium"),
                       colClasses = c("factor", "NULL", "numeric"))


# Display structure of hotdogs2
str(hotdogs2)
```

# readr & data.table

Instead of using read.csv this exercise teaches us how to use read_csv, which is contained in the readr package. The important difference is that instead of a data frame we get a tibble, which is similar to a data frame, but with extended functionality.

```{r}

# Load the readr package
library(readr)

# Import potatoes.csv with read_csv(): potatoes
potatoes <- read_csv('potatoes.csv')
```

The readr equivalent of read.delim is read_tsv.
```{r}
# readr is already loaded

# Column names
properties <- c("area", "temp", "size", "storage", "method",
                "texture", "flavor", "moistness")

# Import potatoes.txt: potatoes
potatoes <- read_tsv('potatoes.txt', col_names=properties)

# Call head() on potatoes
head(potatoes)
```

Confusingly read_delim is the readr version of read.table and is used to read arbitrarily separated data. The parameter delim specifies the separator.
```{r}
properties <- c("area", "temp", "size", "storage", "method",
                "texture", "flavor", "moistness")

# Import potatoes.txt using read_delim(): potatoes
potatoes <- read_delim('potatoes.txt', delim="\t", col_names = properties)


# Print out potatoes
potatoes
```
The parameter skip is used to specify the number of lines that are ignored before reading in data and n_max can be set to limit the number of observations that get imported. For example skip=2, n_max=4, then lines 3, 4, 5, 6 are imported. 
```{r}
# Column names
properties <- c("area", "temp", "size", "storage", "method",
                "texture", "flavor", "moistness")

# Import 5 observations from potatoes.txt: potatoes_fragment
potatoes_fragment <- read_tsv("potatoes.txt", skip = 6, n_max = 5, col_names = properties)
```
col_types is used to specify the types of the columns, c stands for character, d for double, i for integer l for logical and _ skips the column. In the example below all columns contain characters.
```{r}
# Import all data, but force all columns to be character: potatoes_char
potatoes_char <- read_tsv("potatoes.txt", col_types = "cccccccc", col_names = properties)
```

This exercise teaches us to use collector function, which are also used to specify the column types. In the example below the first column is set to be of type factor with the levels Beef, Meat and Poultry. The remaining two columns contain integers.
```{r}
fac <- col_factor(levels = c("Beef", "Meat", "Poultry"))
int <- col_integer()

# Edit the col_types argument to import the data correctly: hotdogs_factor
hotdogs_factor <- read_tsv("hotdogs.txt",
                           col_names = c("type", "calories", "sodium"),
                           col_types = list(fac,int,int))

# Display the summary of hotdogs_factor
summary(hotdogs_factor)
```

The function fread from data.table is similar to the previously discussed functions to read in data, but with one big advantage: it is very fast and the best choice when importing large amounts of data.
```{r}
# load the data.table package
library(data.table)

# Import potatoes.csv with fread(): potatoes
potatoes <- fread("potatoes.csv")
```
The parameters drop and select can be used to drop or select variables from the data. The function plot is used to, you guessed it, create a scatter plot of moistness over texture. 
```{r}
potatoes <- fread("potatoes.csv",select= c(6,8))

# Plot texture (x) and moistness (y) of potatoes
plot( potatoes$texture,   potatoes$moistness)
```

The package readxl is used to load data from xlsx and xls files.
The function excel_sheets lists all the sheets from the excel file specified.
```{r}
# Load the readxl package
library(readxl)

# Print the names of all worksheets
excel_sheets("urbanpop.xlsx")
```

After figuring out which sheets we want to import, the next thing to do is to use read_excel to actually import the data. By setting the sheet parameter we can specify which sheets should be imported. In the example below three sheets are read and then combined to a list.
```{r}
# Read the sheets, one by one
pop_1 <- read_excel("urbanpop.xlsx", sheet = 1)
pop_2 <- read_excel("urbanpop.xlsx", sheet = 2)
pop_3 <- read_excel("urbanpop.xlsx", sheet = 3)

# Put pop_1, pop_2 and pop_3 in a list: pop_list
pop_list <- list(pop_1,pop_2, pop_3)

# Display the structure of pop_list
str(pop_list)
```

Loading in every sheet one-by-one is exhausting. lapply automates that procedure and imports all sheets with one line of code.
```{r}
pop_list <- lapply(excel_sheets("urbanpop.xlsx"), read_excel, path = "urbanpop.xlsx")
# Display the structure of pop_list
str(pop_list)
```

The parameter col_names is used to specify the header, i.e., the column names, for the data frame. It can also be set to TRUE which results in the first line being read as header information or FALSE, in which case read_excel comes up with column names on its own.
```{r}
cols <- c("country", paste0("year_", 1960:1966))
pop_b <- read_excel("urbanpop_nonames.xlsx", col_names = cols)
```

Like with readr, the skip parameter can be specified to skip the first couple of lines of a certain sheet.
```{r}
urbanpop_sel <- read_excel("urbanpop.xlsx",col_names = FALSE,sheet=2, skip = 21)
```

The gdata package extends R with many data manipulation functions. It is another option for reading in .xls files. 
```{r}
library(gdata)

# Import the second sheet of urbanpop.xls: urban_pop
urban_pop <- read.xls("urbanpop.xls", sheet=2)

# Print the first 11 observations using head()
head(urban_pop, n = 11)
```
Due to the underlying implementation of read.xls based on read.csv enables us to use all the previously discussed read.csv options for read.xls, like stringsAsFactors, col.names, skip and many more.
```{r}
columns <- c("country", paste0("year_", 1967:1974))

# Finish the read.xls call
urban_pop <- read.xls("urbanpop.xls", sheet = 2,
                      skip = 50, header = FALSE, stringsAsFactors = FALSE,
                      col.names = columns)

# Print first 10 observation of urban_pop
head(urban_pop, n=10)
```
In this exercise cbind is used to combine the first three sheets of urbanpop.xls, omitting the first columns in sheet 2 and 3. Then na.omit is used to remove all observations that contain NA values. Then, a summary of the combined and cleaned data is printed with the function summary.
```{r}
path <- "urbanpop.xls"
urban_sheet1 <- read.xls(path, sheet = 1, stringsAsFactors = FALSE)
urban_sheet2 <- read.xls(path, sheet = 2, stringsAsFactors = FALSE)
urban_sheet3 <- read.xls(path, sheet = 3, stringsAsFactors = FALSE)

# Extend the cbind() call to include urban_sheet3: urban
urban <- cbind(urban_sheet1, urban_sheet2[-1], urban_sheet3[-1])

# Remove all rows with NAs from urban: urban_clean
urban_clean <- na.omit(urban)
```


# Reproducible Excel work with XLConnect

XLConnect is used as bridge between an excel file and R. In contrast to the previously discussed approaches, the excel file is changed when the data is altered in R. In contrast to readxl and gdata, XLConnect establishes an interface to the file and edits your files from inside R.
```{r}
# urbanpop.xlsx is available in your working directory

# Load the XLConnect package
library("XLConnect")

# Build connection to urbanpop.xlsx: my_book
my_book = loadWorkbook("urbanpop.xlsx")

# Print out the class of my_book
class(my_book)
```
getSheets is used to list all sheets from the excel file, while the function readWorksheet can then be used to read in the specified sheet.
```{r}
my_book <- loadWorkbook("urbanpop.xlsx")

# List the sheets in my_book
getSheets(my_book)

# Import the second sheet in my_book
my_book = readWorksheet(my_book, sheet=2)
```

The parameters startCol and endCol can be set to specify which columns to import. Cbind is used in the example below to combine sheets.
```{r}
# Build connection to urbanpop.xlsx
my_book <- loadWorkbook("urbanpop.xlsx")

# Import columns 3, 4, and 5 from second sheet in my_book: urbanpop_sel
urbanpop_sel <- readWorksheet(my_book, sheet = 2, startCol=3, endCol=5)

# Import first column from second sheet in my_book: countries
countries = readWorksheet(my_book, sheet=2, startCol=1, endCol=1)

# cbind() urbanpop_sel and countries together: selection
selection = cbind(countries, urbanpop_sel)
```

Using the function createSheet we can create new sheets in the excel file. This would not have been possible with the previous approaches.
```{r}
# Build connection to urbanpop.xlsx
my_book <- loadWorkbook("urbanpop.xlsx")

# Add a worksheet to my_book, named "data_summary"
createSheet(my_book, "data_summary")

# Use getSheets() on my_book
getSheets(my_book)
```

This exercise teaches us how to write to a sheet of a Worksheet object. Then we save the current Worksheet to a file called summary.xlsx.
```{r}
# Create data frame: summ
sheets <- getSheets(my_book)[1:3]
dims <- sapply(sheets, function(x) dim(readWorksheet(my_book, sheet = x)), USE.NAMES = FALSE)
summ <- data.frame(sheets = sheets,
                   nrows = dims[1, ],
                   ncols = dims[2, ])

# Add data in summ to "data_summary" sheet
writeWorksheet(my_book, summ, "data_summary")

# Save workbook as summary.xlsx
saveWorkbook(my_book, "summary.xlsx")
```

The function renameSheet can be used to rename sheets. In the example below the sheet data_summary of my_book is renamed to summary.
```{r}
renameSheet(my_book, "data_summary", "summary")
saveWorkbook(my_book, "renamed.xlsx")
```

The function removeSheet is used to remove a sheet from the specified workbook. The sheet can be either determined by giving an index or the name of the sheet, like below.
```{r}
removeSheet(my_book, "summary")
saveWorkbook(my_book, "clean.xlsx")
```
