---
title: "Assignment 4"
author: "Adrian Bracher (Matr. Nr. 01637180)"
date: "21.04.2021"
output:
  pdf_document: 
    toc: true 
    toc_depth: 3
    number_sections: true
---
```{r include=FALSE}
library(dplyr)
library(ggplot2)
counties = readRDS("counties.rds")
babynames = readRDS("babynames.rds")
```

# Transforming Data with dplyr
## Selecting columns
We select the columns state, county, population and poverty with a pipe syntax like below.
```{r}
# Select the columns 
counties %>%
  select(state, county, population, poverty)
```

## Arranging observations
In this exercise we sort the selected columns in descending order on the column public_work.
```{r}
counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

# Add a verb to sort in descending order of public_work
counties_selected %>%
  arrange(desc(public_work))
  
```

## Filtering for conditions
Filter() can be used like below to filter the data using one (or more) criteria.
```{r}
counties_selected <- counties %>%
  select(state, county, population)

# Filter for counties in the state of California that have a population above 1000000
counties_selected %>%
  filter(state == "California", population > 1000000)
```

## Filtering and arranging
The functions arrange() and filter() can also be used consecutively.
```{r}
counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

# Filter for Texas and more than 10000 people; sort in descending order of private_work
counties_selected %>%
  arrange(desc(private_work)) %>%
  filter(state == "Texas", population > 10000)
```

## Calculating the number of government employees
In this exercise we learn that mutate can be used to introduce a new column to the data frame and also that it can be filled with a calculation using the data in other columns.
```{r}
counties_selected <- counties %>%
  select(state, county, population, public_work)

# Sort in descending order of the public_workers column
counties_selected %>%
  mutate(public_workers = public_work * population / 100) %>%
  arrange(desc(public_workers))
```

## Calculating the percentage of women in a county

```{r}
# Select the columns state, county, population, men, and women
counties_selected <- counties %>%
    select(state, county, population, men, women)
  

# Calculate proportion_women as the fraction of the population made up of women
counties_selected %>%
    mutate(proportion_women = women/population)
  
```

## Select, mutate, filter and arrange
In this exercise we combine everything from this chapter.
```{r}
counties %>%
  # Select the five columns 
  select(state, county, population, men, women) %>%
  # Add the proportion_men variable
  mutate(proportion_men = men/population) %>%
  # Filter for population of at least 10,000
  filter(population > 10000) %>%
  # Arrange proportion of men in descending order 
  arrange(desc(proportion_men)) 
```

# Aggregating Data
## Counting by region
Count() is used here to count the number of regions.
```{r eval=FALSE}
# Use count to find the number of counties in each region
counties_selected %>%
  count(region, sort=TRUE)
```

## Counting citizens by state
A weight can be applied too by specifying the parameter wt of the function count.

```{r}
counties_selected <- counties
```


```{r}
# Find number of counties per state, weighted by citizens
counties_selected %>%
  count(state, wt=citizens, sort = TRUE)
```

## Mutating and counting
Mutate and count can be combined for advanced counting with weights.
```{r}
counties_selected %>%
  # Add population_walk containing the total number of people who walk to work 
  mutate(population_walk = population*walk/100) %>%
  count(state, wt=population_walk, sort=TRUE)
```

## Summarizing
Aggregations can be made with the function summarizing, for example calculating the min, max and mean like below.
```{r}
# Summarize to find minimum population, maximum unemployment, and average income
counties_selected %>%
  summarize(min_population = min(population), max_unemployment=max(unemployment), average_income=mean(income))
```

## Summarizing by state
In the exercise below we calculate the population density for each state.
```{r}
# Add a density column, then sort in descending order
counties_selected %>%
  group_by(state) %>%
  summarize(total_area = sum(land_area),
            total_population = sum(population)) %>%
  mutate(density = total_population/total_area) %>%
    arrange(desc(density))
```

## Summarizing by state and region
In this exercise we learn that we can group by multiple columns at once.
```{r}
# Calculate the average_pop and median_pop columns 
counties_selected %>%
  group_by(region, state) %>%
  summarize(total_pop = sum(population)) %>%
  summarize(average_pop = mean(total_pop),
    median_pop = median(total_pop))
```

## Selecting a county from each region
top_n is then used to select the n (the first parameter, in this case 1) rows with the most walking-to-work citizens.

```{r}
# Group by region and find the greatest number of citizens who walk to work
counties_selected %>%
  group_by(region) %>%
  top_n(1, walk)
```

## Finding the highest-income state in each region
In this exercise we select the highest-income state for each region.
```{r}
counties_selected %>%
  group_by(region, state) %>%
  # Calculate average income
  summarize(average_income = mean(income)) %>%
  # Find the highest income state in each region
  top_n(1, average_income)
```

## Using summarize, top_n, and count together
In this exercise we combine everything we've learned in this chapter to count the number of states that have more people living in Metro areas and the ones that have more people living in Nonmetro areas.
```{r}
# Count the states with more people in Metro or Nonmetro areas
counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop = sum(population)) %>%
  top_n(1, total_pop)  %>%
  ungroup()  %>%
  count(metro)
  
```

# Selecting and Transforming Data
## Selecting columns
In the code below different columns are selected and then sorted by the column service.
```{r}
# Glimpse the counties table
glimpse(counties)

counties %>%
  # Select state, county, population, and industry-related columns
  select(state, county, population, professional, service, office, construction, production) %>%
  # Arrange service in descending order 
  arrange(desc(service))
```

## Select helpers
We learn that columns are not only selectable by their exact name, but also with helpers such as ends_with() and starts_with().
```{r}
counties %>%
  # Select the state, county, population, and those ending with "work"
  select(state, county, population, ends_with("work")) %>%
  # Filter for counties that have at least 50% of people engaged in public work
  filter(public_work >= 50)
```

## Renaming a column after count
The function rename can be used to rename a column, for example the default column name n of count to num_counties.
```{r}
# Rename the n column to num_counties
counties %>%
  count(state) %>%
  rename(num_counties = n)
  
```

## Renaming a column as part of a select
Renaming can also be done with select, below we rename the poverty to poverty_rate.
```{r}
# Select state, county, and poverty as poverty_rate
counties %>%
  select(state, county, poverty_rate = poverty)
```

## Using transmute
The verb transmute can be viewed as a combination of the select and mutate  functionality. For an example view the code below.
```{r}
counties %>%
  # Keep the state, county, and populations columns, and add a density column
  transmute(state, county, population, density = population/land_area) %>%
  # Filter for counties with a population greater than one million 
  filter(population > 1000000) %>%
  # Sort density in ascending order 
  arrange(density)
```

## Choosing among the four verbs
There are different use cases for rename, select, mutate and transmute:
- Rename and mutate leave the columns that you don't mention, select and transmute don't.
- Rename and select don't allow calculation, mutate and transmute do.

```{r}
# Change the name of the unemployment column
counties %>%
  rename(unemployment_rate = unemployment)

# Keep the state and county columns, and the columns containing poverty
counties %>%
  select(state, county, contains("poverty"))

# Calculate the fraction_women column without dropping the other columns
counties %>%
  mutate(fraction_women = women / population)

# Keep only the state, county, and employment_rate columns
counties %>%
  transmute(state, county, employment_rate = employed / population)
```

# Case Study: The babynames Dataset
## Filtering and arranging for one year
We use filter and arrange to select rows with year = 1990 and sort them by the number column. This results in the most common names for each year.
```{r}
babynames %>%
  # Filter for the year 1990
  filter(year == 1990) %>%
  # Sort the number column in descending order 
  arrange(desc(number))
```

## Using top_n with babynames
The code below results in the most popular name for each year.
```{r}
# Find the most common name in each year
babynames %>%
    group_by(year) %>%
    top_n(1, number)
```

## Visualizing names with ggplot2
The %in% keywoard checks if a value is found in a vector. Then we plot number over year for the three names selected.
```{r}
# Filter for the names Steven, Thomas, and Matthew 
selected_names <- babynames %>%
  filter(name %in% c("Steven", "Thomas", "Matthew"))

# Plot the names using a different color for each name
ggplot(selected_names, aes(x = year, y = number, color = name)) +
  geom_line()
```

## Finding the year each name is most common 
We combine everything we've learned to find the most common baby name for each year.
```{r}
# Calculate the fraction of people born each year with the same name
babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number)) %>%
  ungroup() %>%
  mutate(fraction = number / year_total) %>%
# Find the year each name is most common
  group_by(name) %>%
  top_n(1, fraction)
```


## Adding the total and maximum for each name
In this exercise we calculate the total occurrences of one name as name_total and also the max occurrence of one name name_max. Then we add a column fraction_max with mutate.
```{r}
 babynames %>%
  group_by(name) %>%
  mutate(name_total = sum(number),
         name_max = max(number)) %>%
  # Ungroup the table 
  ungroup()  %>%
  # Add the fraction_max column containing the number by the name maximum 
  mutate(fraction_max = number/name_max)
```

## Visualizing the normalized change in popularity
In this exercise we filter for three names and then plot the fraction_max column of the previous exercise with ggplot.

```{r include=FALSE}
names_normalized =  babynames %>%
  group_by(name) %>%
  mutate(name_total = sum(number),
         name_max = max(number)) %>%
  # Ungroup the table 
  ungroup()  %>%
  # Add the fraction_max column containing the number by the name maximum 
  mutate(fraction_max = number/name_max)
babynames_fraction <- babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number)) %>%
  ungroup() %>%
  mutate(fraction = number / year_total)
babynames_ratios_filtered <- babynames_fraction %>%
                     arrange(name, year) %>%
                     group_by(name) %>%
                     mutate(ratio = fraction / lag(fraction)) %>%
                     filter(fraction >= 0.00001)
```

```{r}
# Filter for the names Steven, Thomas, and Matthew
names_filtered <- names_normalized %>%
    filter(name %in% c("Steven", "Thomas", "Matthew"))

# Visualize these names over time
ggplot(names_filtered, aes(x=year,y=fraction_max,color=name)) + geom_line()

```

## Using ratios to describe the frequency of a name
In this exercise we learn that the function lag can be used to find the previous value in vector/data frame.
```{r}
babynames_fraction %>%
  # Arrange the data in order of name, then year 
  arrange(name, year) %>%
  # Group the data by name
  group_by(name)  %>%
  # Add a ratio column that contains the ratio between each year 
  mutate(ratio = fraction / lag(fraction))
```

## Biggest jumps in a name
In this exercise we evaluate which names had the biggest jumps in popularity in consecutive years.
```{r}
babynames_ratios_filtered %>%
  # Extract the largest ratio from each name 
  top_n(1, ratio) %>%
  # Sort the ratio column in descending order 
  arrange(desc(ratio))  %>%
  # Filter for fractions greater than or equal to 0.001
  filter(fraction >= 0.001)
```
